---
title: "ggplot_topo_01_country"
author: "Juan Pablo"
date: "2023-04-17"
output: 
  pdf_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R topographic map

### Source:

This is a practice from a video from Milos Popovic

> Author: Milos Popovic

> Youtube video: <https://youtu.be/zoLChBALc1k>

> Github: <https://github.com/milos-agathon/create-crisp-topographic-maps-with-r/blob/main/R/main.r>

### Challenges

-   Use raster images
-   Use `elevatr` to make topographic more realistic with *DEM* data

### Step 1: Setting libraries {#sec-step-1-setting-libraries}

```{r libraries, echo=FALSE}
{
  library(elevatr) # DEM data
  library(giscoR) # geometry data from GISCO
  library(janitor) # cleaning variables names
  library(marmap) # to get a color scale for our relief map
  library(osmdata) # import OpenStreetMap data to crop the raster file
  library(sf) # simple features objects
  library(terra) # to make raster map
  library(tidyverse) # tidy data treatment
}
```

### Step 2: Setting options, constants and other global variables {#sec-step-2-setting-options}

#### Setting up Workspace

```{r echo= FALSE}
#  Cleaning workspace
rm(list = ls())

# Cleaning console
cat("\014")

# gisco cache
gisco_set_cache_dir("./cache/giscoR", install = TRUE)
```

#### Global variables

This variables will make our script more flexible because we can use it for any other country or resolution or another EPSG

```{r}
country <- "IT"
cropping_area <- "locations"
epsg <- "4326"
resolution_geom <- "10"
year_geom <- "2020"
resolution_dem = 7 # from 1 (high-level detail) to 14 (low-level detail)
```

### Step 3: Getting country geometry {#sec-step-3-getting-country-geometry}

We get the country boundary and make a basic plot for testing purposes.

We get this geometry because is needed for `elevatr` to know the limit of the data to retrieve

```{r}
country_geom_raw <- giscoR::gisco_get_countries(
  country = country,
  epsg = epsg,
  resolution = resolution_geom,
  year = year_geom,
)

country_geom <- country_geom_raw |>
  clean_names()

country_geom |>
  ggplot() +
  geom_sf() +
  coord_sf()

```

### Step 4: Getting DEM data for the country

In this case, we are going to use the \`get_elev_raster\` function from from `elevatr` and for the boundaries of the desired country. Currently, this function provides access to the Amazon Web Services Terrian Tiles and the Open Topography global datasets API. A raster object is returned.

Using `get_elev_raster` with `locations` we can add a referenced to crop the global elevation data and this could be a:

-   data frame of x (long) and y (lat) bounding box

-   shape file

-   a sp, sf or raster object

In this case, we will pass the shape file of a country

If we would like to capture the topography of the surrounding countries, then we should define a bounding box in the form of a data frame.

As a `z` argument we used 7 to limit the amount of resources, but we can choose from 1 to 14. Higher the level, higher the resolution.

Finally `clip` is used to crop the data to the passed location

```{r}
get_elevation_data <- function(location, resolution, cropping_area) {
  dem <- elevatr::get_elev_raster(
    clip = cropping_area,
    locations = location,
    z = resolution
  )

  return(dem)
}

data_elevation <- get_elevation_data(country_geom, resolution_dem, cropping_area)
```

```{r}
# testing
terra::plot(data_elevation)
```

```{r}
getwd()
save.image("../working_cache/ggplot_topo_01_country.Rdata")
```

```{r}
load("../working_cache/ggplot_topo_01_country.Rdata")
```

### Credits {#credits}

#### Packages {#packages}

-   elevatr
-   giscoR
-   janitor
-   marmap
-   osmdata
-   sf
-   terra
-   tidyverse
